MCC Kids Model Predictions

### 2019-01-09 stats/kids/kids_model_predictions.Rmd
first draft CogSci paper

---
title: "MCC Kids Model Predictions"
author: "Manuel Bohn"
output: html_document
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(ggthemes)
library(langcog)
library(rwebppl)
library(coda)
library(reshape2)

@@

```
# Pretest

Loading data

```{r pretest data}

pref_data <- read_csv(file="../data/kids_pref_data.csv")%>%
  filter(age != "2",
         trial_type != "train")%>%
  mutate(age_bin = factor(age),
         experiment = "preference_pretest",
         condition = ifelse(change == "false", "same_speaker", "different_speaker"),
         agent = ifelse( change == "false", agent, altAgent))%>%
  select(experiment,subid,age_bin,age_num,trial,condition,agent,correct)

inf_data <- read_csv(file="../data/kids_inf_data.csv")%>%
  filter(subage != "2",
         trial != "filler1",
         trial != "filler2")%>%
  mutate(age_bin = factor(subage),
         experiment = "informativeness_pretest",
         condition = ifelse(control == "no", "test", "control"))%>%
  select(experiment,subid,age_bin,age_num,trial,condition,agent,correct)

pre_data <- bind_rows(pref_data,
                      inf_data)

@@

@@

```{r pretest plot}

pre_p1 <- pre_data %>%
  group_by(experiment,condition,age_num,subid)%>%
  summarise(correct = mean(correct))

ggplot(data = pre_p1, aes(x = age_num, y = correct, col = condition)) +
  geom_smooth(method = "lm", se = F)+
  geom_jitter(alpha = .5, height = 0, width = 0)+
  geom_hline(yintercept = 0.5, lty=2)+
  facet_wrap(~experiment)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  scale_colour_solarized()

@@

@@

```{r rsaUtils, include = FALSE}
rsaUtils <- '
var all_objects = [
{ shape: "triangle", id:1, location: 1},  
{ shape: "triangle", id:2, location: 2},
{ shape: "circle", id:1, location: 2}
]

@@

```{r rsa model, include = FALSE}

rsaModel <- '

@@

var literalListener = cache(function(utterance, priorProbs){
Infer({method: "enumerate", model: function(){
var lexiconName = sample(LexiconPrior); 
var lexicon = lexiconObject[lexiconName];
var obj = sample( Categorical({vs: all_objects, ps: priorProbs}));
if ("label" in utterance) {
 var truthValue = lexicon(utterance, obj);
 condition(truthValue)
}
if (utterance.point) {
 var truthValuePoint = point(utterance, obj);
 condition(truthValuePoint)
}
return obj.shape 
}})
})

@@

## Informativeness

###  Inf prior predictions 
```{r prior speaker optimality slope}
priorSoWebppl <- '

var infData = dataFromR
var priorProbs = [.5, .5, .5] 
var speakerOptimality = []

    map(function(row){

      var speakerOptimality = so_int  + so_slope * (age - row.minage)

      var rsaPredictions = pragmaticListener({label: "dax", point: 2 }, priorProbs, speakerOptimality)

      //display(JSON.stringify(row.minage))

  return extend({so_int: so_int, so_slope: so_slope})
}
'
```

```{r speaker optimality prior pred}
#so_prior <- readRDS("../saves/so_prior.rds")

so_prior<- webppl(
  program_code = paste(rsaUtils, rsaModel, priorSoWebppl , sep='\n'),
  data = inf_data%>%mutate(minage = min(age_num)), 
  data_var = "dataFromR",
  model_var = "model",
  chains = 1,
  inference_opts = list(method = "forward", samples = 1000, verbose = T)
)

saveRDS(so_prior, "../saves/so_prior.rds")

@@

```{r plot prior pred}
ggplot(data = priorPred) +
  geom_abline(aes(intercept = so_int, slope = so_slope), col = "grey", alpha = 0.3)+
  geom_abline(data = mean_so_pre, aes(intercept = int, slope = slope), size = 1)+
  scale_x_continuous(limits=c(0, 2), name="age", breaks = c(0:2), labels = c(3:5)) +
  scale_y_continuous(limits=c(-4, 4), name="Speaker optimality")+
  theme_few()
```

```{r}
ggplot(data = so_prior, aes(x = value, fill = chain)) +
  geom_density(alpha = .5)+
  facet_wrap(~Parameter)+
  theme_few()
```

### Inf posterior predictions

```{r posterior speaker optimality slope}
postSoWebppl <- '

var infData = dataFromR
var priorProbs = [.5, .5, .5]

    map(function(row){

      var rsaPredictions = pragmaticListener({label: "dax", point: 2 }, priorProbs, speakerOptimality)

      //display(rsaPredictions.score( row.correct))
      //display(JSON.stringify(row.correct))
      //display(JSON.stringify(rsaPredictions))

  return extend({so_int: so_int, so_slope: so_slope})
}
'
```

```{r speaker optimality posterior pred}
# so_post <- readRDS("../saves/so_post.rds")

so_post <- webppl(
  program_code = paste(rsaUtils, rsaModel, postSoWebppl , sep='\n'),
  data = inf_data%>%mutate(minage = min(age_num)), 
  data_var = "dataFromR",
  model_var = "model",
  chains = 3,
  cores = 3,
  inference_opts = list(method = "MCMC", samples = 2000, burn = 200, verbose = T)
)

# saveRDS(so_post, "../saves/so_post.rds")

postPred <- so_post%>%
  spread(Parameter, value)%>%
  mutate(chain = factor(Chain))

so_post <- so_post%>%
  mutate(chain = factor(Chain))

mean_so_post <-postPred%>%
  #group_by(chain)%>%
  summarise(int = mean(so_int),
            slope = mean(so_slope))

# confidence intervals
x_plot <- seq(0, 2, by = 0.1)

ci_so_post <- sapply(1:length(postPred$so_int), function(idx) {
           postPred$so_int[idx] + postPred$so_slope[idx] * x_plot
        })

colnames(ci_so_post) <- 1:length(postPred$so_int)
ci_so_post <- as.data.frame(cbind(x_plot, ci_so_post))
ci_so_post <- melt(ci_so_post, id.vars = "x_plot", variable.name = "iteration", 
        value.name = "y_plot")

ci_so_post <- ci_so_post %>%
  group_by(x_plot)%>%
  summarise(ci_upper = hdi_upper(y_plot),
            ci_lower = hdi_lower(y_plot))

```{r plot posterior pred}

ggplot(data = postPred) +
  geom_abline(aes(intercept = so_int, slope = so_slope), col = "grey", alpha = 0.2)+
  geom_ribbon(data = ci_so_post, aes(x =x_plot, ymin = ci_lower, ymax = ci_upper), fill = "darkgreen", alpha = 0.3) +
  geom_abline(data = mean_so_post, aes(intercept = int, slope = slope), size = 1)+
  scale_x_continuous(limits=c(0,2), name="age", breaks = c(0:2), labels = c(3:5)) +
  scale_y_continuous(limits=c(-2, 4), name="Speaker optimality")+
  #facet_wrap(~chain)+
  theme_few()

```{r}
ggplot(data = so_post ) +
  geom_density(aes(x = value, fill = chain), alpha = .5)+
  facet_wrap(~Parameter)+
  theme_few()
```

## Preference

### Pref prior predictions

```{r prior preference congruent}
priorPrefConWebppl <- '

var prefData = dataFromR
var priorProbs = []

    map(function(row){

      var priorProbs= [1-priorReg, 1-priorReg, priorReg]

      var priorPredictions = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: all_objects, ps: priorProbs}));
      return obj.shape == "circle" ? 1 : 0
      }})

      // observe(priorPredictions, row.correct)

     // display(JSON.stringify(priorReg))

    }, prefData)

  return extend({pref_int: pref_int, pref_slope: pref_slope})
}
'
```

```{r prior preference slope}
priorPrefInconWebppl <- '

var prefData = dataFromR
var priorProbs = []

    map(function(row){

      var priorProbs= [priorReg, priorReg, 1-priorReg]

      var priorPredictions = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: all_objects, ps: priorProbs}));
      return obj.shape == "circle" ? 1 : 0
      }})

      // observe(priorPredictions, row.correct)

     // display(JSON.stringify(priorReg))

    }, prefData)

  return extend({pref_int: pref_int, pref_slope: pref_slope})
}
'
```

```{r pref prior pred}

# congruent condition

# pref_prior_con <- readRDS("../saves/pref_prior_con.rds")

pref_prior_con<- webppl(
  program_code = paste(rsaUtils, rsaModel, priorPrefConWebppl , sep='\n'),
  data = pref_data%>%mutate(minage = min(age_num)), 
  data_var = "dataFromR",
  model_var = "model",
  chains = 1,
  inference_opts = list(method = "forward", samples = 1000, verbose = T)
)

 # saveRDS(pref_prior_con, "../saves/pref_prior_con.rds")

# incongruent condition

#pref_prior_incon <- readRDS("../saves/pref_prior_incon.rds")

pref_prior_incon<- webppl(
  program_code = paste(rsaUtils, rsaModel, priorPrefInconWebppl , sep='\n'),
  data = pref_data%>%mutate(minage = min(age_num)), 
  data_var = "dataFromR",
  model_var = "model",
  chains = 1,
  inference_opts = list(method = "forward", samples = 1000, verbose = T)
)

 #saveRDS(pref_prior_incon, "../saves/pref_prior_incon.rds")

```{r}
# congruent condition
prefPriorPredCon <- pref_prior_con%>%
  spread(Parameter, value)

@@

x_plot <- seq(0, 2, by = 0.1)

model_predictions <- sapply(1:length(prefPriorPredCon$pref_int), function(idx) {
            plogis(prefPriorPredCon$pref_int[idx] + prefPriorPredCon$pref_slope[idx] * x_plot)
        })

colnames(model_predictions) <- 1:length(prefPriorPredCon$pref_int)
plot_pref_prior_con <- as.data.frame(cbind(x_plot, model_predictions))
plot_pref_prior_con <- melt(plot_pref_prior_con, id.vars = "x_plot", variable.name = "iteration", 
        value.name = "y_plot")

@@

# incongruent condition

prefPriorPredIncon <- pref_prior_incon%>%
  spread(Parameter, value)

pref_prior_incon <- pref_prior_incon%>%
  mutate(chain = factor(Chain),
         alignment = "incongruent")

model_predictions <- sapply(1:length(prefPriorPredIncon$pref_int), function(idx) {
            plogis(prefPriorPredIncon$pref_int[idx] + prefPriorPredIncon$pref_slope[idx] * x_plot)
        })

colnames(model_predictions) <- 1:length(prefPriorPredIncon$pref_int)
plot_pref_prior_incon <- as.data.frame(cbind(x_plot, model_predictions))
plot_pref_prior_incon <- melt(plot_pref_prior_incon, id.vars = "x_plot", variable.name = "iteration", 
        value.name = "y_plot")

# joining datasets

plot_pref_prior <- bind_rows(
  plot_pref_prior_con %>% mutate(alignment = "congruent"),
  plot_pref_prior_incon %>% mutate(alignment = "incongruent")
)

plot_prior_mean <- plot_pref_prior%>%
  group_by(alignment,x_plot)%>%
  summarise(mean = mean(y_plot),
            ci_upper = hdi_upper(y_plot),
            ci_lower = hdi_lower(y_plot))

pref_prior <- bind_rows(
  pref_prior_con,
  pref_prior_incon
)

```{r plot prior pred}

ggplot(plot_pref_prior) + 
  geom_line(aes(x_plot, y_plot, group = iteration), col = "grey",alpha = 0.2) +
  #geom_ribbon(data = plot_prior_mean, aes(x =x_plot, ymin = ci_lower, ymax = ci_upper), fill = "red", alpha = 0.5) +
  geom_line(data = plot_prior_mean, aes(x_plot, mean), col = "black", size =1) +
  xlab("age") + ylab("p(y | x)") +
  facet_wrap(~alignment)+
  scale_x_continuous(limits=c(0,2), name="age", breaks = c(0:2), labels = c(3:5)) +
  scale_y_continuous(limits = c(0, 1), name="P in favor of preferred object")+
  theme_few()

```{r}
ggplot(data = pref_prior, aes(x = value, fill = chain)) +
  geom_density(alpha = 0.3)+
  facet_grid(Parameter~ alignment)+
  theme_few()
```

### Pref posterior predictions

```{r posterior preference congruent}
postPrefConWebppl <- '

var prefData = dataFromR
var priorProbs = []

    map(function(row){

      var priorProbs= [1-priorReg,1-priorReg, priorReg]

      var priorPredictions = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: all_objects, ps: priorProbs}));
      return obj.shape == "circle" ? 1 : 0
      }})

     // display(JSON.stringify(priorReg))

    }, prefData)

  return extend({pref_int: pref_int, pref_slope: pref_slope})
}
'
```

```{r}
postPrefInconWebppl <- '

var prefData = dataFromR
var priorProbs = []

    map(function(row){

      var priorProbs= [priorReg, priorReg, 1-priorReg]

      var priorPredictions = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: all_objects, ps: priorProbs}));
      return obj.shape == "circle" ? 1 : 0
      }})

     // display(JSON.stringify(priorReg))

    }, prefData)

  return extend({pref_int: pref_int, pref_slope: pref_slope})
}
'
```

```{r pref post pred}

# Same Speaker
# congruent
#pref_post_con_same <- readRDS("../saves/pref_post_con_same.rds")

pref_post_con_same2<- webppl(
  program_code = paste(rsaUtils, rsaModel, postPrefConWebppl , sep='\n'),
  data = pref_data%>%filter(condition == "same_speaker")%>%mutate(minage = min(age_num)), 
  data_var = "dataFromR",
  model_var = "model",
  chains = 3,
  cores = 3,
  inference_opts = list(method = "MCMC", samples = 2000, burn = 200, verbose = T)
)

saveRDS(pref_post_con_same, "../saves/pref_post_con_same.rds")

# Same Speaker
# incongruent
#pref_post_incon_same <- readRDS("../saves/pref_post_incon_same.rds")

pref_post_incon_same<- webppl(
  program_code = paste(rsaUtils, rsaModel, postPrefInconWebppl , sep='\n'),
  data = pref_data%>%filter(condition == "same_speaker")%>%mutate(minage = min(age_num)), 
  data_var = "dataFromR",
  model_var = "model",
  chains = 3,
  cores = 3,
  inference_opts = list(method = "MCMC", samples = 2000, burn = 200, verbose = T)
)

saveRDS(pref_post_incon_same, "../saves/pref_post_incon_same.rds")

# Different Speaker
# congruent
#pref_post_con_diff <- readRDS("../saves/pref_post_con_diff.rds")

pref_post_con_diff<- webppl(
  program_code = paste(rsaUtils, rsaModel, postPrefConWebppl , sep='\n'),
  data = pref_data%>%filter(condition == "different_speaker")%>%mutate(minage = min(age_num)), 
  data_var = "dataFromR",
  model_var = "model",
  chains = 3,
  cores = 3,
  inference_opts = list(method = "MCMC", samples = 2000, burn = 200, verbose = T)
)

 saveRDS(pref_post_con_diff, "../saves/pref_post_con_diff.rds")

# Different Speaker
# incongruent
#pref_post_incon_diff <- readRDS("../saves/pref_post_incon_diff.rds")

pref_post_incon_diff<- webppl(
  program_code = paste(rsaUtils, rsaModel, postPrefInconWebppl , sep='\n'),
  data = pref_data%>%filter(condition == "different_speaker")%>%mutate(minage = min(age_num)), 
  data_var = "dataFromR",
  model_var = "model",
  chains = 3,
  cores = 3,
  inference_opts = list(method = "MCMC", samples = 2000, burn = 200, verbose = T)
)

saveRDS(pref_post_incon_diff, "../saves/pref_post_incon_diff.rds")

```{r}
# Same Speaker
# congruent

prefPostPredConSame <- pref_post_con_same%>%
  spread(Parameter, value)

pref_post_con_same <- pref_post_con_same%>%
  mutate(chain = factor(Chain))%>%
  mutate(alignment = "congruent",
         condition = "same speaker")

x_plot <- seq(0, 2, by = 0.1)

post_predictions_con_same <- sapply(1:length(prefPostPredConSame$pref_int), function(idx) {
            plogis(prefPostPredConSame$pref_int[idx] + prefPostPredConSame$pref_slope[idx] * x_plot)
        })

colnames(post_predictions_con_same) <- 1:length(prefPostPredConSame$pref_int)
plot_post_data_con_same <- as.data.frame(cbind(x_plot, post_predictions_con_same))
plot_post_data_con_same <- melt(plot_post_data_con_same, id.vars = "x_plot", variable.name = "iteration", 
        value.name = "y_plot")%>%
  mutate(alignment = "congruent",
         condition = "same speaker")

# Same Speaker
# incongruent

prefPostPredInconSame <- pref_post_incon_same%>%
  spread(Parameter, value)

pref_post_incon_same <- pref_post_incon_same%>%
  mutate(chain = factor(Chain))%>%
  mutate(alignment = "incongruent",
         condition = "same speaker")

post_predictions_incon_same <- sapply(1:length(prefPostPredInconSame$pref_int), function(idx) {
            plogis(prefPostPredInconSame$pref_int[idx] + prefPostPredInconSame$pref_slope[idx] * x_plot)
        })

colnames(post_predictions_incon_same) <- 1:length(prefPostPredInconSame$pref_int)
plot_post_data_incon_same <- as.data.frame(cbind(x_plot, post_predictions_incon_same))
plot_post_data_incon_same <- melt(plot_post_data_incon_same, id.vars = "x_plot", variable.name = "iteration", 
        value.name = "y_plot")%>%
  mutate(alignment = "incongruent",
         condition = "same speaker")

# Different speaker
# congruent
prefPostPredConDiff <- pref_post_con_diff%>%
  spread(Parameter, value)

pref_post_con_diff <- pref_post_con_diff%>%
  mutate(chain = factor(Chain))%>%
  mutate(alignment = "congruent",
         condition = "different speaker")

post_predictions_con_diff <- sapply(1:length(prefPostPredConDiff$pref_int), function(idx) {
            plogis(prefPostPredConDiff$pref_int[idx] + prefPostPredConDiff$pref_slope[idx] * x_plot)
        })

colnames(post_predictions_con_diff) <- 1:length(prefPostPredConDiff$pref_int)
plot_post_data_con_diff <- as.data.frame(cbind(x_plot, post_predictions_con_diff))
plot_post_data_con_diff <- melt(plot_post_data_con_diff, id.vars = "x_plot", variable.name = "iteration", 
        value.name = "y_plot")%>%
  mutate(alignment = "congruent",
         condition = "different speaker")

# Different speaker
# incongruent
prefPostPredInconDiff <- pref_post_incon_diff%>%
  spread(Parameter, value)

pref_post_incon_diff <- pref_post_incon_diff%>%
  mutate(chain = factor(Chain))%>%
  mutate(alignment = "incongruent",
         condition = "different speaker")

post_predictions_incon_diff <- sapply(1:length(prefPostPredInconDiff$pref_int), function(idx) {
            plogis(prefPostPredInconDiff$pref_int[idx] + prefPostPredInconDiff$pref_slope[idx] * x_plot)
        })

colnames(post_predictions_incon_diff) <- 1:length(prefPostPredInconDiff$pref_int)
plot_post_data_incon_diff <- as.data.frame(cbind(x_plot, post_predictions_incon_diff))
plot_post_data_incon_diff <- melt(plot_post_data_incon_diff, id.vars = "x_plot", variable.name = "iteration", 
        value.name = "y_plot")%>%
  mutate(alignment = "incongruent",
         condition = "different speaker")

# combining datasets

pref_post <- bind_rows(
  pref_post_con_diff,
  pref_post_incon_diff,
  pref_post_con_same,
  pref_post_incon_same
)

plot_post_data <- bind_rows(
  plot_post_data_incon_same,
  plot_post_data_con_same,
  plot_post_data_incon_diff,
  plot_post_data_con_diff
)

plot_post_data_mean <- plot_post_data%>%
  group_by(alignment,condition,x_plot)%>%
  summarise(mean = mean(y_plot),
            ci_upper = hdi_upper(y_plot),
            ci_lower = hdi_lower(y_plot))
```

ggplot(plot_post_data) + 
  geom_line(aes(x_plot, y_plot, group = iteration), col = "grey", alpha = 0.2) + 
  geom_ribbon(data = plot_post_data_mean, aes(x =x_plot, ymin = ci_lower, ymax = ci_upper), fill = "darkgreen", alpha = 0.3) +
  geom_line(data = plot_post_data_mean, aes(x_plot, mean), col = "black") + 
  scale_x_continuous(limits=c(0,2), name="age", breaks = c(0:2), labels = c(3:5)) +
  scale_y_continuous(limits = c(0, 1), name="P in favor of preferred object")+
  facet_grid(alignment~condition)+
  theme_few()

```{r}
ggplot(data = pref_post, aes(x = value, fill = chain)) +
  geom_density(alpha = 0.3)+
  facet_grid(Parameter ~ alignment*condition)+
  theme_few()
```

# 2 vs object world

```{r posterior preference congruent}
postPrefConWebppl2 <- '

var prefData = dataFromR
var priorProbs = []

    map(function(row){

     // display(JSON.stringify(priorReg))

    }, prefData)

  return extend({pref_int: pref_int, pref_slope: pref_slope})
}
'
```

# Same Speaker
# congruent
#pref_post_con_same <- readRDS("../saves/pref_post_con_same.rds")

pref_post_con_same2<- webppl(
  program_code = paste(rsaUtils, rsaModel, postPrefConWebppl2 , sep='\n'),
  data = pref_data%>%filter(condition == "same_speaker")%>%mutate(minage = min(age_num)), 
  data_var = "dataFromR",
  model_var = "model",
  chains = 3,
  cores = 3,
  inference_opts = list(method = "MCMC", samples = 2000, burn = 200, verbose = T)
)

prefPostPredConSame2 <- pref_post_con_same2%>%
  spread(Parameter, value)

pref_post_con_same2 <- pref_post_con_same2%>%
  mutate(chain = factor(Chain))%>%
  mutate(alignment = "congruent",
         condition = "same speaker")

x_plot <- seq(0, 2, by = 0.1)

post_predictions_con_same2 <- sapply(1:length(prefPostPredConSame2$pref_int), function(idx) {
            plogis(prefPostPredConSame2$pref_int[idx] + prefPostPredConSame2$pref_slope[idx] * x_plot)
        })

colnames(post_predictions_con_same2) <- 1:length(prefPostPredConSame2$pref_int)
plot_post_data_con_same2 <- as.data.frame(cbind(x_plot, post_predictions_con_same2))
plot_post_data_con_same2 <- melt(plot_post_data_con_same2, id.vars = "x_plot", variable.name = "iteration", 
        value.name = "y_plot")%>%
  mutate(alignment = "congruent",
         condition = "same speaker",
         world = "2 objects")

x <- bind_rows(
  plot_post_data_con_same2,
  plot_post_data_con_same
)

x_mean <- x%>%
  group_by(world,x_plot)%>%
  summarise(mean = mean(y_plot),
            ci_upper = hdi_upper(y_plot),
            ci_lower = hdi_lower(y_plot))

a <- prefPostPredConSame2%>%
  summarise(mean_int = mean(pref_int),
            mean_slope = mean(pref_slope))%>%
  expand(y = plogis(mean_int + mean_slope * x_plot))%>%
  mutate(x = x_plot,
         world = "2 objects")

b <- prefPostPredConSame%>%
  summarise(mean_int = mean(pref_int),
            mean_slope = mean(pref_slope))%>%
  expand(y = plogis(mean_int + mean_slope * x_plot))%>%
  mutate(x = x_plot,
         world = "3 objects")

c <- bind_rows(a,b)

ggplot(x) + 
  geom_line(aes(x_plot, y_plot, group = iteration), col = "grey", alpha = 0.2) + 
  geom_ribbon(data = x_mean, aes(x =x_plot, ymin = ci_lower, ymax = ci_upper), fill = "darkgreen", alpha = 0.3) +
  geom_line(data = x_mean, aes(x_plot, mean), col = "black") + 
  geom_line(data = c, aes(x, y), col = "red") + 
  scale_x_continuous(limits=c(0,2), name="age", breaks = c(0:2), labels = c(3:5)) +
  scale_y_continuous(limits = c(0, 1), name="P in favor of preferred object")+
  facet_grid(~world)+
  theme_few()

@@

@@

var levels = function(df, label){
  return _.uniq(_.map(df, label));
}
var prefData = filter(dataFromR, {experiment: "preference"})
var infData = filter(dataFromR, {experiment: "informativeness"})

// make finer grained as needed
var binned_age_range = _.range(0, 2, 0.1)

@@

//var priorProbs = []

@@

    var conditionData = filter(prefData, {condition: cndtn})

@@

     // display(JSON.stringify(priorReg))

@@

  var diff_speaker_params = filter(pref_params, 
        {condition: "different_speaker"})[0]

  var same_speaker_params = filter(pref_params, 
        {condition: "same_speaker"})[0]

@@

    return extend(
    //diff_speaker_params, same_speaker_params, 
{
     // diff_priorReg,
    //  same_priorReg,
    //  speakerOptimality,
      rsaPredictions_diff_cong,
      rsaPredictions_diff_incong,
      rsaPredictions_same_cong,
      rsaPredictions_same_incong,
      age_bin,
    //  so_slope,
    //  so_int
    })

@@

>>>>

---
title: "MCC Kids Model Predictions"
author: "Manuel Bohn"
output: 
  html_document:
      code_folding: hide
---

```{r, include = F}
knitr::opts_chunk$set(echo = TRUE, warning = F)
```

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(ggthemes)
library(ggpubr)
library(langcog)
library(rwebppl)
library(coda)
library(reshape2)

@@

@@

# Pretest

```{r pretest data, include=FALSE}
pref_data <- read_csv(file="../data/kids_pref_data.csv")%>%
  filter(age != "2",
         trial_type != "train")%>%
  mutate(age_bin = factor(age),
         experiment = "preference_pretest",
         condition = ifelse(change == "false", "same_speaker", "different_speaker"),
         agent = ifelse( change == "false", agent, altAgent),
         minage = min(age_num))%>%
  select(experiment,subid,age_bin,age_num,trial,condition,agent,correct,minage)

inf_data <- read_csv(file="../data/kids_inf_data.csv")%>%
  filter(subage != "2",
         trial != "filler1",
         trial != "filler2")%>%
  mutate(age_bin = factor(subage),
         experiment = "informativeness_pretest",
         condition = ifelse(control == "no", "test", "control"),
        minage = min(age_num))%>%
  select(experiment,subid,age_bin,age_num,trial,condition,agent,correct,minage)

pre_data <- bind_rows(pref_data,
                      inf_data)
```

@@

This is the data we use to estimate the speaker optimality nad the age specific prior.

```{r pretest plot so, echo = F}

pre_so <- pre_data %>%
  filter(experiment == "informativeness_pretest")%>%
  group_by(age_num,subid)%>%
  summarise(correct = mean(correct))

plot_inf <- ggplot(data = pre_so, aes(x = age_num, y = correct)) +
  geom_jitter(alpha = 1, height = 0.1, width = 0.1, col = "grey")+
  geom_smooth(method = "lm", se = T, col = "black",fill = "darkgreen", alpha = .4, size = .4)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="age",y="Proportion Expected Choice")+
  theme_few() +
  ggtitle("Informativeness data")+
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  coord_fixed(ratio = 2)
```

```{r pretest plot prior, echo = F}

pre_prior <- pre_data %>%
  filter(experiment == "preference_pretest")%>%
  group_by(condition,age_num,subid)%>%
  summarise(correct = mean(correct))

plot_pref <- ggplot(data = pre_prior, aes(x = age_num, y = correct, col = condition)) +
  geom_jitter(alpha = 0.5, height = 0.1, width = 0.1)+
  geom_smooth(aes(fill = condition),method = "lm", se = T, col = "black",  alpha = .5, size = .4)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  ggtitle("Preference data")+
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  coord_fixed(ratio = 2)+
  scale_color_solarized()+
  scale_fill_solarized()

@@

```{r, echo = F, fig.width= 8}
ggarrange(plot_inf, plot_pref, widths = c(1.29,2))
```

@@

### Model utensils.

There are two types of objects, one used in the 3-object world of the RSA model (`all_objects`) and one used in the 2-object world of the prior (`prior_objects`).

```{r rsaUtils}
rsaUtils <- '
var all_objects = [
{ shape: "triangle", id:1, location: 1},  
{ shape: "triangle", id:2, location: 2},
{ shape: "circle", id:1, location: 2}
]

@@

### RSA model code

```{r rsa model}

rsaModel <- '
var literalListener = cache(function(utterance, priorProbs){
Infer({method: "enumerate", model: function(){
var lexiconName = sample(LexiconPrior); 
var lexicon = lexiconObject[lexiconName];
var obj = sample( Categorical({vs: all_objects, ps: priorProbs}));
if ("label" in utterance) {
 var truthValue = lexicon(utterance, obj);
 condition(truthValue)
}
if (utterance.point) {
 var truthValuePoint = point(utterance, obj);
 condition(truthValuePoint)
}
return obj.shape 
}})
})

@@

@@

### Priors for intercept and slope for speaker optimality and preference

Sanity check for the priors on slope and intercept.

#### Speaker optimality

```{r prior speaker optimality slope}
priorSoWebppl <- '
var infData = dataFromR
var priorProbs = [.5, .5, .5] 
var speakerOptimality = []
var model  = function(){
  var so_slope = uniformDrift({a: -2, b: 2, width: 0.4})
  var so_int = uniformDrift({a: -2, b: 2, width: 0.4})
    map(function(row){
      var age = row.age_num
      var speakerOptimality = so_int  + so_slope * (age - row.minage)
      var rsaPredictions = pragmaticListener({label: "dax", point: 2 }, priorProbs, speakerOptimality)

      //display(JSON.stringify(row.minage))
    }, infData)
  return extend({so_int: so_int, so_slope: so_slope})
}
'
```

```{r speaker optimality prior pred}
so_prior <- readRDS("../saves/so_prior.rds")
# so_prior<- webppl(
#   program_code = paste(rsaUtils, rsaModel, priorSoWebppl , sep='\n'),
#   data = inf_data%>%mutate(minage = min(age_num)), 
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 1,
#   inference_opts = list(method = "forward", samples = 1000, verbose = T)
# )
#saveRDS(so_prior, "../saves/so_prior.rds")

@@

so_prior_plot <- ggplot(data = priorPred) +
  geom_abline(aes(intercept = so_int, slope = so_slope), col = "grey", alpha = 0.3)+
  geom_abline(data = mean_so_pre, aes(intercept = int, slope = slope), size = 0.4)+
  scale_x_continuous(limits=c(0, 2), name="age", breaks = c(0:2), labels = c(3:5)) +
  scale_y_continuous(limits=c(-4, 4), name="Speaker optimality")+
  ggtitle("Model speaker optimality prior")+
  coord_fixed(ratio=1/4)+
  theme_few()
```

#### Preference

```{r prior preference congruent}
priorPreWebppl <- '
var prefData = dataFromR
var priorProbs = [] 
var logistic = function(x) {1 / (1 + Math.exp(-x))}
var model  = function(){
  var pref_slope = uniformDrift({a: -2, b: 2, width: 0.4})
  var pref_int = uniformDrift({a: -2, b: 2, width: 0.4})
    map(function(row){
      var age = row.age_num
      var priorReg = logistic(pref_int + pref_slope * (age - row.minage))
      var priorProbs= [1-priorReg, priorReg]
      var priorPredictions = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: prior_objects, ps: priorProbs}));
      return obj.shape == "circle" ? 1 : 0
      }})

      // observe(priorPredictions, row.correct)
     // display(JSON.stringify(priorReg))
    }, prefData)
  return extend({pref_int: pref_int, pref_slope: pref_slope})
}
'
```

```{r}
pref_prior_con <- readRDS("../saves/pref_prior_con.rds")

# pref_prior_con<- webppl(
#   program_code = paste(rsaUtils, rsaModel, priorPrefConWebppl , sep='\n'),
#   data = pref_data%>%mutate(minage = min(age_num)), 
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 1,
#   inference_opts = list(method = "forward", samples = 1000, verbose = T)
# )
 # saveRDS(pref_prior_con, "../saves/pref_prior_con.rds")

prefPriorPredCon <- pref_prior_con%>%
  spread(Parameter, value)

@@

x_plot <- seq(0, 2, by = 0.1)
model_predictions <- sapply(1:length(prefPriorPredCon$pref_int), function(idx) {
            plogis(prefPriorPredCon$pref_int[idx] + prefPriorPredCon$pref_slope[idx] * x_plot)
        })
colnames(model_predictions) <- 1:length(prefPriorPredCon$pref_int)
plot_pref_prior_con <- as.data.frame(cbind(x_plot, model_predictions))
plot_pref_prior_con <- melt(plot_pref_prior_con, id.vars = "x_plot", variable.name = "iteration", 
        value.name = "y_plot")

@@

pref_prior_plot <- ggplot(plot_pref_prior_con) + 
  geom_line(aes(x_plot, y_plot, group = iteration), col = "grey",alpha = 0.2) +
  #geom_ribbon(data = plot_prior_mean, aes(x =x_plot, ymin = ci_lower, ymax = ci_upper), fill = "red", alpha = 0.5) +
  geom_line(data = plot_prior_con_mean, aes(x_plot, y_plot), col = "black", size = 0.4) +
  xlab("age") + ylab("Proportion Expected Choice") +
  scale_x_continuous(limits=c(0,2), name="age", breaks = c(0:2), labels = c(3:5)) +
  scale_y_continuous(limits = c(0, 1), name="P in favor of preferred object")+
  ggtitle("Model preference prior")+
  coord_fixed(ratio=2)+
  theme_few()
```

```{r, echo = F, fig.width= 8}
ggarrange(so_prior_plot, pref_prior_plot)
```

There is no information in the priors. All structure in the models further down comes from the data

@@

## Model code for predictions

@@

var levels = function(df, label){
  return _.uniq(_.map(df, label));
}
var prefData = _.filter(dataFromR, {experiment: "preference_pretest"})
var infData = _.filter(dataFromR, {experiment: "informativeness_pretest"})

// make finer grained as needed
var binned_age_range = _.range(0, 2, 0.05)

@@

@@

    var conditionData = _.filter(prefData, {condition: cndtn})

@@

@@

  var diff_speaker_params = _.filter(pref_params, 
        {condition: "different_speaker"})[0]

  var same_speaker_params = _.filter(pref_params, 
        {condition: "same_speaker"})[0]

@@

    return extend(
    // diff_speaker_params, same_speaker_params, 
      {
      diff_priorReg,
      same_priorReg,
      speakerOptimality,
      rsaPredictions_diff_cong,
      rsaPredictions_diff_incong,
      rsaPredictions_same_cong,
      rsaPredictions_same_incong,
      age_bin,
      so_slope,
      so_int,

    })

@@

```{r, echo = F, fig.width= 8}
ggarrange(pred_so_plot, plot_inf)
```

Here we estimate the speaker optimality parameter that best captures the data for the different ages. The model based estimate nicely captures the developmental trend.

```{r plot prior model pred, echo = F}
plot_model_pred_prior <- kids_model_pred_prior%>%
  group_by(age_bin,condition)%>%
  summarise(mean = mean(prop_preferred),
            lci = hdi_lower(prop_preferred),
            uci = hdi_upper(prop_preferred))
